---
title: "Rising Ventures: Exploring Unicorn Insights"
output:
  html_document: default
  word_document: default
  pdf_document: default

---




## Summary

‘Unicorn’ is a term used in the venture capital industry to describe private startup companies valued at over $1 billion.
The project aims to conduct an in-depth analysis of Unicorn startups to uncover insights regarding their relationships with the respective industries, valuation trends, and the years they entered the market. This analysis seeks to unveil trends within Unicorn industries and understand their potential implications on valuation. Additionally, it intends to identify key investors and evaluate their potential influence on the respective industries.




## Data

The dataset employed for this project comprises seven columns: Company, Valuation ($B), Date Joined (indicating the date when the company became a unicorn), Country, City, Industry, and Investors.




## Methodology

To approach the project objectives, the analysis is divided in:

1. Industry Trends Analysis: Explore and identify prevailing trends within Unicorn industries. Assess how factors like the country, the city and the valuation impact these industries.Analyse the most prevalent industries in each city where unicorns thrive, the primary cities serving as hubs for unicorn companies and distribution of unicorns based on countries.
	
2. Valuation Dynamics: Examine if and how industry trends correlate with the valuation of Unicorn startups. Identify  the valuation range observed among unicorn companies, and whether certain industries or specific trends tend to result in higher valuations. 
	
3. Investor Impact Assessment: Investigate the involvement of investors in these industries. Identify patterns, and the subsequent impact on industry growth throughout the years. Which industries witness higher investor interest and investment allocation.
	
4. Historical Analysis: Conduct a historical analysis of the entrance of Unicorns into various industries. Determine if certain periods or industry developments coincide with an influx of successful startups. Examine correlation between years and the emergence of unicorn companies.", 




## Loading packages and data preparation


```{r}
library(dplyr)
library(ggplot2)
library(plotly)
library(tidyr)
library(ggpubr)
library(treemap)

```


Loading the dataset


```{r}
unicorns <- read.csv('C:/Users/ikry/OneDrive/Desktop/MSc DS/CODING/R/r project unicorn/unicorns_dataset.csv')
head(unicorns)
```
Data cleaning


Despite the dataset was already cleaned, some additional cleaning was necessary. For instance, I consolidated certain industries that were initially categorized separately due to syntax errors. The date column was transformed into a "Year Joined" column as the primary focus lies on the year variable. Furthermore, I converted the valuation column into a numeric format to enable mathematical manipulations and analysis.


```{r}
unicorns$Valuation...B. <- as.numeric(gsub("[\\$,]", "", unicorns$Valuation...B.))

unicorns <- unicorns %>%
  mutate(Industry = case_when(
    Industry == 'Artificial intelligence' ~ 'Artificial Intelligence',
    Industry == 'Internet' ~ 'Internet software & services',
    TRUE ~ Industry  
  ))

unicorns <- unicorns %>%
  mutate(Date.Joined = as.Date(Date.Joined, format = "%m/%d/%Y")) %>%  
  mutate(Year.Joined = format(Date.Joined, "%Y")) %>%  
  select(-Date.Joined)
```

## 1.Industries Analysis

```{r}
industry_counts <- unicorns %>%
  group_by(Industry) %>%
  summarise(count = n()) %>%
  filter(count >= 2) # only industries with at least 2 startups

head(industry_counts, 15)

# Industries distribution bar plot
industries_distribution <-ggplot(industry_counts, aes(x = Industry, y = count, fill = Industry)) +
  geom_bar(stat = "identity") +
  labs(x = "Industry", y = "Number of Startups") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

industries_distribution
```

Top 5 unicorn countries

```{r}
country_counts <- table(unicorns$Country)

top5 <- head(sort(country_counts, decreasing = TRUE), 5)
head(top5) #Tab1.1

top5_startups <- subset(unicorns, Country %in% c('United States', 'China', 'India', 'United Kingdom', 'Germany'))
head(top5_startups)

top5_countries_evaluation <- top5_startups %>%
  group_by(Country) %>%
  summarise(Total_Valuation = sum(as.numeric(`Valuation...B.`)), .groups = 'drop')
head(top5_countries_evaluation) #Tab1.2


```


### 1.1 Unicorn Cities


```{r}

city_counts <- top5_startups %>% 
  group_by(City.) %>%
  summarise(Count = n()) %>%
  filter(Count > 10) %>%
  arrange(desc(Count))



fig1.2 <- ggplot(city_counts, aes(x = reorder(City., Count), y = Count)) +
  geom_bar(stat = "identity", fill = "skyblue", color = "black") +
  geom_text(aes(label = Count), vjust = -0.5, size = 3, color = "black") +  
  labs(title = "Cities of Unicorns",
       x = "City",
       y = "Number of Unicorns") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

fig1.2
```

### 1.2 Most Popular Industry per City


```{r}


#what are the most popular industries per city
city_counts <- top5_startups %>%
  group_by(City.) %>%
  summarise(Count = n()) %>%
  filter(Count > 10) %>%
  arrange(desc(Count))

top_cities <- city_counts$City.[1:10]

# Filter the dataframe to include only the top cities
df_top_cities <- unicorns %>%
  filter(City. %in% top_cities)

city_industry_counts <- df_top_cities %>%
  group_by(City., Industry) %>%
  summarise(Count = n()) %>%
  ungroup()

#  most popular industry for each city
most_popular_industry <- city_industry_counts %>%
  arrange(City., desc(Count)) %>%
  group_by(City.) %>%
  slice(1) %>%
  ungroup()

head(most_popular_industry, 15)

fig1.3 <- ggplot(most_popular_industry, aes(x = City., y = Count, fill = Industry)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Most Popular Industries per City",
       x = "City",
       y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_brewer(palette = "Set3") 

fig1.3 
```

```{r}
selected_cities <- c("San Francisco", "New York", "Beijing", "Shangai", "London")  


subset_df <- subset(city_industry_counts, City. %in% selected_cities)


top10_cities <- city_industry_counts[city_industry_counts$City. %in% selected_cities, ]

fig1.4 <- treemap(top10_cities, index = c("City.", "Industry"), vSize = "Count",
        title = "Industries for Each City", palette = "RdYlBu")  


```




The top five countries hosting the highest number of startups are the USA, China, India, the United Kingdom, and Germany, as depicted in Tab1.1, Fig1.1 illustrates the respective valuations of these countries' startups.
The visualization on cities by the number of unicorns (Fig1.2) highlights San Francisco and New York as the primary hubs not only in the USA but also globally. Following  are two Chinese cities, Beijing and Shanghai, followed by the UK's capital city. Bengaluru in India is in the fifth position.

Notably, there's a considerable difference among the top three cities, with almost a 50-unit disparity. However, the distribution becomes relatively more even among the subsequent cities. It's important to note that only cities in the USA host more than 100 startups, signaling their dominance in the market.
In the graph (Fig1.3) depicting "Popular Industries per City," the data highlights the prevalence of the Internet Software and Services industry as the most prominent sector across the top five countries. Where San Francisco exhibits dominance compared to other cities, standing out significantly without any close competitors or rivals in this sector. Following closely behind are industries related to artificial intelligence, and health occupies the third position in terms of popularity among these cities.

The treemap (Fig1.4) provides a more comprehensive insight into the distribution of industries across the different cities.



## 2. Valuation Dynamics


```{r}
sum_by_year <- unicorns %>%
  group_by(Year.Joined) %>%
  summarise(Total_Valuation = sum(`Valuation...B.`))

# AMOUNT OF INVESTEMENTS BY YEAR
Fig2.1 <- ggplot(sum_by_year, aes(x = Year.Joined, y = Total_Valuation)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(x = "Year", y = "Total Valuation ($B)",
       title = "Total Valuation by Year") +
  theme_minimal()

Fig2.1



#VALUATIONS OF TOP5 COUNTRIES
total_valuation_by_country <- top5_startups %>%
  group_by(Country, Year.Joined) %>%
  summarise(Total_Valuation = sum(`Valuation...B.`, na.rm = TRUE))
total_valuation_by_country #Fig2.2
```


```{r}
#Some industry statistics
industry_stats <- unicorns %>%
  group_by(Industry) %>%
  summarise(
    Mean_Valuation = mean(Valuation...B., na.rm = TRUE),
    Median_Valuation = median(Valuation...B., na.rm = TRUE)
  ) %>%
  arrange(desc(Mean_Valuation))  

print(industry_stats)

fig2.3 <- ggplot(unicorns, aes(x = Industry, y = Valuation...B.)) +
  geom_boxplot() +
  labs(x = "Industry", y = "Valuation") +
  ggtitle("Distribution of Valuations across Industries")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size=8)) 

fig2.3 


```


The visualization depicting the "Total Valuation per Year" (Fig2.1) showcases a consistent upward trend in valuations over the past five years, culminating in a peak in 2021 with remarkably high valuations. However, the data indicates a return to the distribution pattern observed in the preceding years by the year 2022.


Fig 2.2 elucidates that the notable peak observed in 2021 primarily stemmed from a substantial increase in valuation within the USA. Contrasting with other top 5 countries whose valuations ranged between 0 and 100 in 2021, the USA's valuation soared significantly, surpassing 600. This disparity in valuation notably contributed to the overall peak observed in 2021.


## 3.Stakeholders analysis


```{r}

investor <- unicorns %>%
  separate_rows(Investors, sep = ',\\s*') %>%
  group_by(Investor = trimws(Investors)) %>%
  summarise(Count = n())


# Filter out rows where Investor is null or empty
investors <- investor %>%
  filter(!is.na(Investor) & Investor != "") #tab 3.1



mean_investments <- mean(investors$Count)
cat("The average number of investments made by investors is:", round(mean_investments, 2))



#filter out only those who invested more than 10 
investor_counts <- investors %>%
  filter(Count > 10)


head(investor_counts) #tab 3.2


investors <- investor_counts %>%
  plot_ly(
    x = ~Investor,
    y = ~Count,
    size = ~Count,
    color = ~Investor,
    type = 'scatter',
    mode = 'markers'
  )

investors 
```

### 3.1 Where do investor invest?

```{r}

industry_valuation <- top5_startups %>%
  group_by(Industry) %>%
  summarise(Total_Valuation = sum(`Valuation...B.`), .groups = 'drop')



fig3 <- ggplot(industry_valuation, aes(y = Industry, x = Total_Valuation)) +
  geom_bar(stat = "identity", fill = "lightgreen") +
  labs(title = "Industry-wise Valuation", y = "Industry", x = "Valuation ($B)")

fig3
```
 
 
In the investor table (Tab3.1), an initial count revealed a total of 1343 investors. On average, these investors have invested in approximately 2,5 startups.

Upon applying a filter to include only those who invested in more than 10 startups, the count reduced to only 42 investors. (Tab 3.2)

The industry-wise valuation plot illustrates the preferred industries by investors, showcasing Fintech as the most dominant with a valuation surpassing 700 billion dollars. Following Fintech, Internet services hold a substantial position, while e-commerce and artificial intelligence share a similar valuation. Conversely, the travel industry exhibits the least valuation among these industries. (Fig3)


## 4.Hystorical Analysis

when unicorns started appearing in cities throughout the years?


```{r}
city_counts <- top5_startups %>%
  group_by(City.) %>%
  summarise(Count = n()) %>%
  filter(Count > 10) %>%
  arrange(desc(Count))

top_cities <- city_counts$City.[1:10]

#  include only the top cities
df_top_cities <- unicorns %>%
  filter(City. %in% top_cities)

city_year_counts <- df_top_cities %>%
 count(City., Year.Joined)


fig4.1 <- plot_ly(city_year_counts, x = ~Year.Joined, y = ~n, color = ~City., type = 'scatter', mode = 'lines+markers') %>%
  layout(title = "Unicorn appearances by City",
         xaxis = list(title = "Year"),
         yaxis = list(title = "Number of Startups"))


fig4.1 
```

## 4.1 Industries Growth by Year


```{r}

yearly_country_valuation <- top5_startups %>%
  group_by(Year.Joined, Country) %>%
  summarise(Total_Valuation = sum(`Valuation...B.`, na.rm = TRUE)) %>%
  ungroup()


print(yearly_country_valuation)

fig4.2 <- plot_ly(yearly_country_valuation, x = ~Year.Joined, y = ~Total_Valuation, color = ~Country, type = 'scatter', mode = 'lines+markers') %>%
  layout(title = "Total Valuation by Year and Country",
         xaxis = list(title = "Year"),
         yaxis = list(title = "Total Valuation"))


fig4.2 

```


Although nearly all countries experienced an almost threefold increase in total valuation from their initial year entering the market to the subsequent year (with the exception of the UK, which witnessed a decrease, potentially influenced by its initially high valuation), the USA showcased a notable pattern. Beginning at USD 5.7 billion in 2011, the USA rapidly escalated to USD 156 billion in 2012. Subsequently, they exhibited fluctuating patterns, alternating between notably high valuations and more modest ones in the following years.

When examining valuation growth in China, it's evident that it experienced an upward trajectory from 2014 to 2017, reaching a peak aggregate valuation of USD 185.8 billion. However, this trend reversed in the subsequent years until 2020, with a slight uptick noted in 2021, reaching USD 82.0 billion. Conversely, India's unicorn aggregate valuation displayed a more consistent pattern compared to the fluctuations observed in the United States and China. It demonstrated a steady incline throughout the entire period from 2014 to 2021, ultimately reaching its pinnacle of USD 73.4 billion in 2021.






## 4.2  Valuation-Year Joined Realtionship  


```{r}
covariance_val_year <- cov(unicorns$Valuation, as.numeric(substring(unicorns$Year.Joined, 1, 4)))

correlation_val_year <- cor(unicorns$Valuation, as.numeric(substring(unicorns$Year.Joined, 1, 4)))


print(paste("Correlation:", correlation_val_year))
print(paste("Covariance:", covariance_val_year))
```

The correlation coefficient of approximately -0.23 indicates a weak negative linear relationship between 'Valuation' and 'Year.Joined'.
The covariance of approximately -3.50 indicates that there is a negative linear relationship, but the magnitude doesn't provide as clear an interpretation as correlation, especially because covariance is influenced by the scale of the variables.


While it's important to note that correlation and covariance solely measure linear relationships between variables and do not imply causation, it's reasonable to anticipate that certain industries, such as artificial intelligence, might have made significant market entrances in the recent past.



Is there any potential correlations between industry and the Year joined?

```{r}

fig4.3 <- plot_ly(data = top5_startups, x = ~Year.Joined, y = ~Industry, type = "histogram2d") %>%
  layout(
    title = "Date Joined by Unicorns",
    titlefont = list(size = 20),
    paper_bgcolor = '#edeeee',
    plot_bgcolor = '#edeeee',
    hovermode = 'closest'
    
  )

fig4.3  

```




Figure 4.1 illustrates the emergence of unicorns in various cities over the years….

Meanwhile, Figure 4.2 focuses on demonstrating the fluctuations in industries' valuations over the course of different years.

Based on the information derived from the dataset, it appears there is a weak negative linear relationship between the valuation of startups and the year they became unicorns. The correlation coefficient of approximately -0.23 signifies this weak negative linear correlation. 

Additionally, the covariance of approximately -3.50 also suggests a negative linear relationship between these variables. However, the magnitude of covariance doesn't offer as straightforward an interpretation as correlation, mainly because covariance is influenced by the scales of the variables involved.

Subsequently, I began exploring whether a relationship existed between the year of entrance in the market and the various industries represented. I plotted a heatmap (Fig 4.3) to visually represent any potential correlations. Interestingly, the heatmap revealed that the year 2021 held significant importance for both the fintech and internet industries, and they maintained a similar but diminished result in 2022.




## Conclusion


This analysis unravels a dynamic and thriving landscape within the startup domain, particularly evident in specific countries and cities. San Francisco and New York emerge as primary global hubs, followed by Beijing, Shanghai, London, and Bengaluru. The Internet Software and Services industry dominates, notably in San Francisco registering a consistent upward trend in valuations over five years, peaking in 2021.

The standout valuation surge experienced by the USA in 2021, particularly in Fintech and internet industries, continues into 2022, albeit at a diminished level. By previous analysis in the industries we can say that San Francisco and New York played a pivotal role as depicted in Fig 4.1.
Additionally, the return to previous valuation patterns in 2022 aligns with the weak negative correlation observed between startup valuation and the year of unicorn inception, offering intriguing insights into the pivotal year of 2021 for specific industries.

By analyzing the industry growth over the past 5 years, it becomes evident that sectors such as Fintech and Internet Software & Services have undergone remarkable expansion, particularly in the years 2021 and 2022 as depicted in fig. 4.3. Conversely, other industries have displayed a more consistent growth trend, reaching their peak in 2021 in most instances. 2018 stood out as one of the most prosperous year for unicorn emergence in terms of valuation, primarily associated with industries such as E-commerce & Direct-to-consumer,Fintech, and Internet Software & Services.

Interestingly, from the initial count of 1343 investors;the majority is a one-time investor and only the 3.13% (42) invest in more than 10 startups. The preferred industries are Fintech, followed by Internet services, e-commerce, and AI. The travel industry has the least valuation.

Based on the analysis of investor, Accel stands out as the most successful company, having invested in 62 unicorns to date. Following closely in this ranking are 'Andreessen Horowitz' and 'Tiger Global Management' with investments in 59 and 56 unicorns respectively.

To sum up, based on the table 'top5_countries_evaluation'(tab 1.2), the United States firmly maintains its global unicorn market dominance with an aggregate valuation of USD 2,000 billion representing the 64% of the market. Following behind is China, with a valuation of almost USD 700 billion  commanding roughly 21% of the market. India, with a valuation of USD 203 billion holds about 6% of the market share. Just like United Kingdom even though India leads in the overall number of unicorns.Germany, with 80,88 USD billion total evaluation represents the 2,5% of the unicorn market.






















